---
title: "R and RStudio Tips & Tricks"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Shortcuts

> "Name a keyboard shortcut in RStudio that changed your life"

https://twitter.com/kc_analytics/status/1378770803454529538?s=20

- ALT + SHIFT + K = Open shortcut overview

- CTRL + Shift + M = %>% 
- CTRL + ALT + I = Insert code chunk
- CTRL + SHIFT + R = Insert section header (in script)
- ALT + - = <-

- ALT + UP/DOWN = move code line up or down
- ALT + SHIFT + UP/DOWN = copy current line and move up/down (instead of copy & paste)

- CTRL + . = Go to file/function (especially useful to find function definitions in packages/big projects)
- TAB = Auto complete (really useful for file paths when reading files, e.g., within `read_csv("path/to/file.csv")``)

# Visualization

Start branding your graphs
https://michaeltoth.me/you-need-to-start-branding-your-graphs-heres-how-with-ggplot.html
- define custom ggplot theme
- define custom color palette (depending on project)

# Vectorizing

Stop using loops and start using `purrr::map_...`

- often way faster
- less code
- specify output type

Example: read 10 same-structured files and row-bind them

```{r, eval=FALSE}
# loop-style
library(tidyverse)

df <- tibble()

for (i in 1:length(vector)) {
  df_new <- read_csv(files[i])
  df <- bind_rows(df, df_new)
}
```

```{r, eval=FALSE}
# purrr-style
library(tidyverse)

df <- purrr::map_dfr(files, read_csv)
```

- `vectorize()`
- capture side effects: `possibly()`, `safely()`
- iteratively (depend on previously computed values)
- `accumulate()`

# Reading files

Consider different methods of reading/writing/storing data

- `base::read.csv()` (standard)
- `base::save()`, `base::saveRDS()` (.Rdata, .rds)
- `readr::read_csv()` (faster, tidyverse)
- `data.table::fread()` (even faster)
- `vroom::vroom()` (even faster, but still a bit experimental)
- `arrow::read_arrow()` (.parquet, smaller size than .csv)

# Twitter

- Follow `#rstats`
- Follow RStudio employees

# R independently from RStudio (Till)

- Use R with VIM editor
- RStudio VIM editor emulation
- pre-defined templates (e.g. for YAML headers)
- even more (personalizable) short cuts

# ggplot2 & plotly (Dimitri)

- Interactive HTML ggplots

```{r, eval=FALSE}
library(ggplot2)
library(plotly)

p <- ggplot2(aes(...)) +
  ...
ggplotly(p)
```

- `ggforce` doesn't work with `plotly`

# RStudio Add-ins (Haseeb)

- point and click addin for ggplot
  - `ggplotAssist`
  - `esquisse`
- `styler`: nice-looking code

# Fonts with code-specific ligatures (Haseeb)

- ligature support (e.g. `<-`)
favorite fonts:
- jetbeans mono (jetbrains)
- fira code

# Tips to make regex less painful

- https://regexr.com/
- https://regex101.com/
- https://github.com/gadenbuie/regexplain

# Bring tables into a data frame/tibble (Janine)

- `datapasta` addin
- copy and paste (web) tables into tibbles

# Citations in R Markdown

- `citr` addin (pre-installed in RStudio >= 1.4)

# Tidy time series (machine learning) (Haseeb)

- packages `modeltime`, `timetk`
- by https://twitter.com/mdancho84

# Tidytuesday

- https://github.com/rfordatascience/tidytuesday
- Twitter: `#tidytuesday`
- Live-recorded screen casts https://twitter.com/drob


# Ideas for future meetups

- Loops vs. Vectorization: "Bring your loop and we're vectorizing it"
- Regex
- RStudio Addins